Retrofitting Effect Handlers onto OCaml

### 2020-11-17 retro-concurrency.tex
edits

Fibers are allocated on the heap using |malloc| and are |free|d when no longer
necessary. We do not use a \emph{stack cache} of recently freed stacks for
speeding up fiber allocation~\cite{Farvardin20}. In our experiments, we
observed that directly using |mimalloc|~\cite{Leijen19} outperforms a stack
cache.

@@

Next, we have the variable-sized area for the OCaml frames. In order to keep
fibers small, this area is initially 16 words in length. When the stack pointer
|rsp| becomes less than the stack threshold (maintained in the |Caml_state|
table), the stack is said to have overflown. On stack overflow, we copy the
whole fiber to a new area with double the size. In Multicore OCaml, we
introduce stack overflow checks into the function prologue of OCaml functions.
In our evaluation of real world OCaml programs, we observed that most function
calls are to leaf functions with small frame size. In order to take advantage
of this fact, we introduce a small red zone at the top of the stack. The
compiler elides the stack overflow check for leaf functions whose frame size is
less than the size of the red zone.

@@

We also present micro benchmarks results in Table~\ref{tab:micro_noeffect}.
Since micro benchmarks magnify micro-architectural optimisations, we also
report the number of instructions executed (obtained using |perf|) along with
time. |exnval| perform 100 million iterations of installing exception handlers
and returning with value. |exnraise| is similar, but raises an exception in
each iteration. |extcall| and |callback| perform 100 million external calls and
callbacks to identity functions. The other micro benchmarks are highly
recursive programs and were taken from ~\cite{Farvardin20}. For micro benchmarks,
we observed that padding tightloops with a few |nop| instructions, which
changes the loop alignment, makes the code upto 15\% faster. Hence, the
difference in running times under 15\% may not be statistically significant.

>>>>

Fibers are allocated on the heap using |malloc| and are |free|d when no longer
necessary. We do not use a \emph{stack cache} of recently freed stacks for
speeding up fiber allocation~\cite{Farvardin20} as we
observed in our evaluation (\S\ref{sec:eval}) that directly using |mimalloc|~\cite{Leijen19} outperforms a stack cache.

@@

Next, we have the variable-sized area for the OCaml frames, known as the red zone.
In order to keep fibers small, this area is initially 16 words in length.
When the stack pointer |rsp| becomes less than the stack threshold
(maintained in the |Caml_state| table), the stack is said to have overflown.
On stack overflow, we copy the
whole fiber to a new area with double the size. In Multicore OCaml, we
introduce stack overflow checks into the function prologue of OCaml functions.
In our evaluation of real world OCaml programs ((\S\ref{sec:eval}) we observed that most function
calls are to leaf functions with a small frame size. We introduce a small red zone at the top of the stack to take advantage of this fact. The
compiler elides the stack overflow check for leaf functions whose frame size is
less than the size of the red zone.

@@

We also present micro benchmarks results in Table~\ref{tab:micro_noeffect}.
Since micro benchmarks magnify micro-architectural optimisations, we also
report the number of instructions executed (obtained using |perf|) along with
time. |exnval| perform 100 million iterations of installing exception handlers
and returning with value. |exnraise| is similar, but raises an exception in
each iteration. |extcall| and |callback| perform 100 million external calls and
callbacks to identity functions. The other micro benchmarks are highly
recursive programs and were taken from ~\cite{Farvardin20}. For micro benchmarks,
we observed that padding tightloops with a few |nop| instructions, which
changes the loop alignment, makes the code up to 15\% faster. Hence, the
difference in running times under 15\% may not be statistically significant.
